On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-09-06 Manuscript/manuscript.tex
Fixed typos and errors

The objective to this paper is to analyze the reaction times at the 2022 World 
Track and Field Championships to check for any abnormalities, and we approach
this in two different ways.  First we model the data, which are
the reaction times of every non-disqualification since the 1999 World 
Championships, using a gamma linear mixed-effects model with two random effects 
one effect for the year and a second effect for the
heat within each year.
Based on the fitted model, we can then estimate the probability of a reaction 
time being below a certain threshold (e.g., 0.1). 
Second, we compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship.  By
comparing reaction times from the same athlete across multiple competitions we 
are able to determine if the 2022 reaction times differed significantly from 
those of other competitions.
Our investigation into reaction times is important not only because of its
implications for future races, but also because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times, which is the goal of
every sprinter who steps onto the track.
\eds{Since the World Championships are happening again now, is data from this 
year already available? If it is available, we would need to justify its omission
in the current analysis.
If it is not available, but becomes available while under review, would we 
reanalyze the data?  Does it seem likely our conclusions would change based on
what is going on this year?}

@@

There are two types of data that are explored throughout this paper: data from
1999-2023 for every world championship, and data from athletes who competed in
the 2022 World Championships and another competition (either 2019 World 
Championships, 2023 World Championships or a 2022 national competition typically
held within three months of the 2022 World Championships). 
%\eds{is this true? are you still using the 2019 world championships data for the
%rank-based analyses, or only the national competitions?}
%\of{Yes the ranked based comparison has 2 subsections: 2022 national vs 2022
%international and 2019 international vs 2022 international.  Should I instead say
%that there are three types of data across 2 sections or something like that?}
%\eds{I think this is fine here, but it is not at all clear that there are two
%subsections in the rank-based results for 2022 national vs 2022
%international and 2019 international vs 2022 international.}
%\of{Currently under the Methods/Results section I have GLMM and rank based comparison
%do you think I need subsections under rank based comparison, or is that too much
%compartmentalization?  Or should I instead say there are 3 types of data that are
%used in two different methods, or something like that?} \eds{I prefer
%subsections within rank-based comparison.}
The first data set is used for the GLMM analyses described in 
Section~\ref{sec:glmm}, while the second collection of data is used for the 
rank-based comparisons described in Section~\ref{sec:rank}.  This paper was
started before the 2023 World Championships, but as that data has become
available recently, we are pleased to include it in our analysis.

\subsection{World Championships 1999--2022}\label{sec:dataworld}

Data was taken from the World Athletics 
and covers the men's 110 meter hurdles from 1999 to 2022; the source can be 
found in the Appendix.
Although women's data is available, we chose not to include women's data in the
analyses to follow due to the numerous papers citing differences in men and
women's reaction times \citep[e.g.,][]{lipps2011implications, 
babicc2009reaction, panoutsakopoulos2020gender}.
For the creation of the GLMM model we focus on the reaction times during the 
semifinal and final heats only, as reaction times from preliminary heats are 
often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}.

@@

From 2007--2009, the IAAF (the former name for World Athletics, and the 
governing body for the World Track and Field Championships) instituted a rule 
change that allowed one warning false start before a sprinter was disqualified 
\citep{iaaf2009falsestart}. There were 18 male and 7 female false starts at the 
2007 World Championships, and 18 male and 7 female false starts at the 2009 World 
Championships. Starting in 2011, the rule was scrapped, and the old policy which
automatically disqualifies runners who false start was reinstated. This was 
desirable for World Athletics as false starts can make already lengthy track 
meets more tedious, both for athletes and viewers watching the television 
broadcast. By returning to the harsher policy and cracking down on false starts,
World Athletics had reduced men's false starts by two thirds in 2011 (6 male and
4 female disqualifications) \citep{iaaf2009falsestart}. \citet{haugen2013effect}
examined the effect of different false start rules that IAAF has imposed from 
1997 to 2009 and found statistically significant improvement in reaction times 
under more lenient rules.

@@

\subsection{GLMM}\label{sec:glmm}
Based on an exploratory analysis using the GLMM model for the reaction times
from the World Championships, the final model we chose was a gamma mixed-effect 
model. As suggested by the log-likelihood and the Akaike Information
Criterion (AIC), the normal error model on the log-scale was found to be
much inferior to the gamma model which better captures the skewness of the
data. Two random effects were found useful: one captures the venue effect in
each year and the other captures the heat effect within each year.
Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
Conditional on a venue effect $v_i$ of year~$i$ and a heat effect
$h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
gamma distribution. 
\eds{I was talking to my husband about this. Could the heat effect be justified 
to capture the variability in the amount of time between the ``set'' and the 
gun going off (assuming this is different from heat to heat)? E.g., quicker 
gunshots after the ``set'' could lead to slower reaction times (or the reverse), 
but this effect would be shared across all athletes in the same heat.}
In the classic generalized linear model (GLM) sense,
$Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
dispersion parameter~$\phi > 0$. The conditional mean is characterized by
\[
g\{E(Y_{ijk} | v_i, h_{i/j})\} = \alpha + v_i + h_{i/j},
\]
where $g$ is a known link function,
$\gamma_i$ is normally distributed with mean zero and
variance~$\sigma_v^2$, and $\delta_{i/j}$ is normally distributed with mean
zero and variance~$\sigma_h^2$.
Although \citet{lo2015idlink} suggested that the gamma 
family with an identity link works well for reaction time in cognitive
psychological research, the logarithm
link function worked better for our data and additionally ensures the positivity
of the conditional mean. The gamma mixed-effect model can be fit with the 
\texttt{glmer()} function in R package \texttt{lme4} \citep{lme4}.

@@

With our small sample size, the asymptotic normal distribution may not be
reliable, so we use 1000 random permutations to simulate the observed test 
statistic: $W^*$ given number of times to obtain the p-value of said test 
statistics. 
\eds{I do not understand what is meant in the text after the colon in the 
previous sentence.  Please rephrase.}
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} and \texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

A summary of the two effect model is shown in Table~\ref{tab:Gamma_parameters}
for the pooled datasets and shows the standard error of the year
effect increases for both the men's final and men's pooled data sets when
including 2022. We choose to only look at the pooled datasets, as R displayed
warnings when creating the finals only models.  The two effect model with venue
and heat effect was the best model, as indicated by an ANOVA test 
\eds{I think it may be better to call this a likelihood ratio test rather than
an ANOVA test, assuming that is the formal test being used} and decrease in
AIC.  For the entire pooled dataset, the AIC for the venue effect only model was
$-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
two effect model, which we will present throughout this section, was $-2075.7$. 
Thus, the two
effect model is superior and will be presented throughout the rest of the paper.
\of{How specific do I need to be in this section? Will people who read this paper
run the code to see the p value or should I say it directly and how do I say it?}
\eds{If you are doing a likelihood ratio test, I suggest including the p-value 
and the degrees of freedom associated with the test.}

\begin{table}
  \centering
  \caption{Summary statistics for the two effect GLMM on various data sets. SE
  is short for standard error. \eds{data in table needs to be revisited and table
	needs to be referenced within the text. Also, it needs to be clarified 
	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}} 
  \begin{tabular}{c c c c c c}
      \toprule
      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\delta_{i/j}$ \\
      \midrule
      Excluding 2022 &  0.057 & 0.033 & $-1.868$ & 0.136 \\
      All Data & $-2075.7$ 0.057 & 0.040 & $-1.880$ & 0.013 \\
      \bottomrule
  \end{tabular}
  \label{tab:Gamma_parameters}
\end{table}

Two aims of this section are to determine how extreme a reaction time below
0.1 seconds is, and to determine how the exclusion of 2022 could alter the results.
Using the gamma function with link set to log, the probability of the reaction 
time being less than 0.1 seconds can be calculated in R using the methods 
described in Section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
chosen to estimate the probability of observing a reaction time below 
two thresholds: 0.1 and 0.08. When examining the All Pooled dataset, a probability
of $5.15\cdot10^{-3}$ corresponds to roughly 1 in every 195 starts
being below 0.1 seconds.  When we exclude 2022, the probability of observing a
reaction time below 0.1 seconds decreased to $4.15\cdot10^{-3}$, or 1 in every
240 starts.
\of{I feel like with the two effect model the probability change is not as
different as it was previously}

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than 0.1 and 0.08
  for various data sets.}
  \begin{tabular}{c c c c} 
   \toprule
   Data Set & P(R. Time $<$ 0.08) & P(R. Time $<$ 0.09) & P(R. Time $<$ 0.1)  \\ 
   \midrule
   Old Pooled & $4.67\cdot10^{-5}$ & $5.91\cdot10^{-4}$ &  $4.15\cdot10^{-3}$  \\ 
   All Pooled & $6.01\cdot10^{-5}$ & $7.39\cdot10^{-4}$ & $5.15\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}

We can use the same model but alter our interpretation slightly to select a 
reaction time barrier given the probability of observing a time less than the 
barrier.  For example, if we want a 0.1\% chance of observing an illegally fast 
reaction time for the men's pooled dataset that includes 2022, then the reaction
time barrier should be 0.089 seconds, as shown in Table~\ref{tab:Sim_time}.  We
can repeat this calculation for each dataset and for various probability levels,
for example a probability of 0.0001 (one disqualification for every 10,000 starts)
results in a reaction time barrier of 0.079 for the men's pooled dataset.  
Table~\ref{tab:Sim_time} shows that the inclusion of 2022 results in a lower
reaction time barrier, further showing the claims above that the reaction times
in 2022 were low.

\begin{table}
  \centering
  \caption{Suggested reaction time barriers based on probability of observing
  a time less than the barrier. \of{Not sure if this makes sense}}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & P(RxnTime $<$ x = $10^{-3}$) & P(RxnTime $<$ x = $10^{-4}$) \\ 
   \midrule
   Old Pooled & $0.093$ & $0.083$ \\ 
   All Pooled & $0.091$ & $0.082$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_time}
\end{table}

@@

Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
\texttt{B = 1000} indicating that the random permutation size was chosen to be 
1000, as discussed previously in Section~\ref{sec:rank}.  
%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
%that is preferred.  
In both cases, the national versus international comparison
produced stronger \eds{lower?} p values than the 2019 versus 2022 comparison, with the
national vs international \texttt{exact = TRUE} test producing a p value of 
$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
is a highly significant test result and shows that athletes who competed at the
2022 Track and Field World Championships were on average, significantly faster
at reacting than they were a few months prior.  For the 2019 versus 2022
same athlete comparison with \texttt{exact = TRUE} argument, the p value was
significant at $\alpha = 0.01$, however when \texttt{exact = FALSE}, the p
value was higher and only significant at $\alpha = 0.05$. 
\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
Nonetheless, these
tests help to show that athletes competing at the 2022 Track and Field World
Championships had faster reaction times than when they had competed previously.

\begin{table}
  \centering
  \caption{}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic \\ 
   \midrule
   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10{-2}$ \\ 
   Nat vs Inter & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. Nat vs Inter compares data
  from the same athletes who competed at the 2022 national-level championships
  and the 2022 World Track and Field Championships.}
  \label{tab:Clusrankresults}
\end{table}

@@

This paper does not consider data from the women's 100 meter hurdle, and although
that was initially because of concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 was at best only
slightly faster relative to other championships going back to 2003 
%\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
%the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
%results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
%womens 100m hurdles.} 
Likewise, the rank-based comparisons resulted in p values
that were not significant. 
\eds{could that be because of even smaller sample sizes?}
  The lack of evidence from the women's data makes it 
harder to conclude that the sensors used at the 2022 championships were faulty, 
although it is unknown if the sensors for the men's 110 meter hurdle and women's
100 meter hurdle were the same.  
\eds{is any calibration of the sensors needed (e.g. for weight?) that might
also play a role for women vs men. (My husband asked me about this as a possible 
reason for the difference in results for men and women, but he did not know
if any calibration is actually done.)}

>>>>

The objective to this paper is to analyze the reaction times at the 2022 World 
Track and Field Championships to check for any abnormalities, and we approach
this in two different ways.  First we model the data, which are
the reaction times of every non-disqualification since the 1999 World 
Championships, using a gamma linear mixed-effects model with two random effects 
one effect for the year and a second effect for the
heat within each year.
Based on the fitted model, we can then estimate the probability of a reaction 
time being below a certain threshold (e.g., 0.1). 
Second, we compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship.  By
comparing reaction times from the same athlete across multiple competitions we 
are able to determine if the 2022 reaction times differed significantly from 
those of other competitions.
Our investigation into reaction times is important not only because of its
implications for future races, but also because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times, which is the goal of
every sprinter who steps onto the track.
\eds{Since the World Championships are happening again now, is data from this 
year already available? If it is available, we would need to justify its omission
in the current analysis.
If it is not available, but becomes available while under review, would we 
reanalyze the data?  Does it seem likely our conclusions would change based on
what is going on this year?}
\of{Resolved}

@@

There are two types of data that are explored throughout this paper: data from
1999-2023 for every world championship, and data from athletes who competed in
the 2022 World Championships and another competition (either 2019 World 
Championships, 2023 World Championships or a 2022 national competition typically
held within three months of the 2022 World Championships). 
%\eds{is this true? are you still using the 2019 world championships data for the
%rank-based analyses, or only the national competitions?}
%\of{Yes the ranked based comparison has 2 subsections: 2022 national vs 2022
%international and 2019 international vs 2022 international.  Should I instead say
%that there are three types of data across 2 sections or something like that?}
%\eds{I think this is fine here, but it is not at all clear that there are two
%subsections in the rank-based results for 2022 national vs 2022
%international and 2019 international vs 2022 international.}
%\of{Currently under the Methods/Results section I have GLMM and rank based comparison
%do you think I need subsections under rank based comparison, or is that too much
%compartmentalization?  Or should I instead say there are 3 types of data that are
%used in two different methods, or something like that?} \eds{I prefer
%subsections within rank-based comparison.}
The first data set is used for the GLMM analysis described in 
Section~\ref{sec:glmm}, while the second collection of data is used for the 
rank-based comparisons described in Section~\ref{sec:rank}.  This paper was
started before the 2023 World Championships, but as that data has become
available recently, we are pleased to include it in our analysis.

\subsection{World Championships 1999--2023}\label{sec:dataworld}

Data was taken from the World Athletics 
and covers the men's 110 meter hurdles from 1999 to 2023; the source can be 
found in the Appendix.
Although women's data is available, we chose not to include women's data in the
analyses to follow due to the numerous papers citing differences in men and
women's reaction times \citep[e.g.,][]{lipps2011implications, 
babicc2009reaction, panoutsakopoulos2020gender}.
For the creation of the GLMM model we focus on the reaction times during the 
semifinal and final heats only, as reaction times from preliminary heats are 
often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}.

@@

From 2007 to 2009, the IAAF (the former name for World Athletics, and the 
governing body for the World Track and Field Championships) instituted a rule 
change that allowed one warning false start before a sprinter was disqualified 
\citep{iaaf2009falsestart}. There were 18 male and 7 female false starts at the 
2007 World Championships, and 18 male and 7 female false starts at the 2009 World 
Championships. Starting in 2011, the rule was scrapped, and the old policy which
automatically disqualifies runners who false start was reinstated. This was 
desirable for World Athletics as false starts can make already lengthy track 
meets more tedious, both for athletes and viewers watching the television 
broadcast. By returning to the harsher policy and cracking down on false starts,
World Athletics had reduced men's false starts by two thirds in 2011 (6 male and
4 female disqualifications) \citep{iaaf2009falsestart}. \citet{haugen2013effect}
examined the effect of different false start rules that IAAF has imposed from 
1997 to 2009 and found statistically significant improvement in reaction times 
under more lenient rules.

@@

\subsection{GLMM}\label{sec:glmm}
Based on an exploratory analysis using the GLMM model for the reaction times
from the World Championships, the final model we chose was a gamma mixed-effect 
model. As suggested by the log-likelihood and the Akaike Information
Criterion (AIC), the normal error model on the log-scale was found to be
much inferior to the gamma model which better captures the skewness of the
data. Two random effects were found useful: one captures the venue effect in
each year and the other captures the heat effect within each year.  The heat
effect is important as it captures the variability in the amount of time
athletes are on the starting blocks before the sun goes off.  Thus, we can try
to reduce this variability by accounting for this in the model through the batch
effect.
Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
Conditional on a venue effect $v_i$ of year~$i$ and a heat effect
$h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
gamma distribution. 
\eds{I was talking to my husband about this. Could the heat effect be justified 
to capture the variability in the amount of time between the ``set'' and the 
gun going off (assuming this is different from heat to heat)? E.g., quicker 
gunshots after the ``set'' could lead to slower reaction times (or the reverse), 
but this effect would be shared across all athletes in the same heat.}
\of{Yes I agree with this and updated the language in the text.}
In the classic generalized linear model (GLM) sense,
$Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
dispersion parameter~$\phi > 0$. The conditional mean is characterized by
\[
g\{E(Y_{ijk} | v_i, h_{i/j})\} = \alpha + v_i + h_{i/j},
\]
where $g$ is a known link function,
$\gamma_i$ is normally distributed with mean zero and
variance~$\sigma_v^2$, and $\delta_{i/j}$ is normally distributed with mean
zero and variance~$\sigma_h^2$.
Although \citet{lo2015idlink} suggested that the gamma 
family with an identity link works well for reaction time in cognitive
psychological research, the logarithm
link function worked better for our data and additionally ensures the positivity
of the conditional mean. The gamma mixed-effect model can be fit with the 
\texttt{glmer()} function in R package \texttt{lme4} \citep{lme4}.

@@

With our small sample size, the asymptotic normal distribution may not be
reliable, so we use 1000 random permutations to simulate the observed test 
statistic.
\eds{I do not understand what is meant in the text after the colon in the 
previous sentence.  Please rephrase.}
\of{I'm not sure what that was, I removed it now.}
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} and \texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

A summary of the two effect model is shown in Table~\ref{tab:Gamma_parameters}
for the pooled datasets and shows the standard error of the year
effect increases when
including 2022.  The two effect model with venue
and heat effect was the best model, as indicated by an ANOVA test 
\eds{I think it may be better to call this a likelihood ratio test rather than
an ANOVA test, assuming that is the formal test being used} and decrease in
AIC.  For the entire data, the AIC for the venue effect only model was
$-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
two effect model, which we will present throughout this section, was $-2075.7$. 
Thus, the two
effect model is superior and will be presented throughout the rest of the paper.
\of{How specific do I need to be in this section? Will people who read this paper
run the code to see the p value or should I say it directly and how do I say it?}
\eds{If you are doing a likelihood ratio test, I suggest including the p-value 
and the degrees of freedom associated with the test.}

\begin{table}
  \centering
  \caption{Summary statistics for the two effect GLMM on various data sets. SE
  is short for standard error. \eds{data in table needs to be revisited and table
	needs to be referenced within the text. Also, it needs to be clarified 
	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}} 
  \begin{tabular}{c c c c c c}
      \toprule
      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\delta_{i/j}$ \\
      \midrule
      Excluding 2022 & $-1912.9$ &  0.058 & 0.033 & $-1.868$ & 0.136 \\
      All Data & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134 \\
      \bottomrule
  \end{tabular}
  \label{tab:Gamma_parameters}
\end{table}

Two aims of this section are to determine how extreme a reaction time below
0.1 seconds is, and to determine how the exclusion of 2022 could alter the results.
Using the gamma function with link set to log, the probability of the reaction 
time being less than 0.1 seconds can be calculated in R using the methods 
described in Section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
chosen to estimate the probability of observing a reaction time below 
two thresholds: 0.1 and 0.08. When including 2022, a probability
of $4.38\cdot10^{-3}$ corresponds to roughly 1 in every 228 starts
being below 0.1 seconds.  When we exclude 2022, the probability of observing a
reaction time below 0.1 seconds decreased to $3.52\cdot10^{-3}$, or 1 in every
285 starts.

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than 0.08, 0.09, and
  0.01 seconds based on the two effect model.}
  \begin{tabular}{c c c c} 
   \toprule
   Data Set & P(R. Time $<$ 0.08) & P(R. Time $<$ 0.09) & P(R. Time $<$ 0.1)  \\ 
   \midrule
   Excluding 2022 & $3.19\cdot10^{-5}$ & $4.48\cdot10^{-4}$ &  $3.52\cdot10^{-3}$  \\ 
   Including 2022 & $4.01\cdot10^{-5}$ & $5.82\cdot10^{-4}$ & $4.38\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}

We can use the same model but alter our interpretation slightly to select a 
reaction time barrier given the probability of observing a time less than the 
barrier.  When considering all of the data, if we want a 0.1\% chance of 
observing an illegally fast reaction time, the reaction
time barrier should be 0.092 seconds, as shown in Table~\ref{tab:Sim_time}.  We
can repeat this calculation for various probability levels,
for example a probability of 0.0001 (one disqualification for every 10,000 starts)
results in a reaction time barrier of 0.083 seconds.  As the exclusion of 2022
from the data resulted in a lower probability of observing a low reaction time,
the reaction time barrier will be less lenient when calculating the barrier
based on probability.

\begin{table}
  \centering
  \caption{Suggested reaction time barriers based on probability of observing
  a time less than the barrier. \of{Not sure if this makes sense}}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & P(RxnTime $<$ x = $10^{-3}$) & P(RxnTime $<$ x = $10^{-4}$) \\ 
   \midrule
   Old Pooled & $0.093$ & $0.084$ \\ 
   All Pooled & $0.092$ & $0.083$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_time}
\end{table}

@@

Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
\texttt{B = 1000} indicating that the random permutation size was chosen to be 
1000, as discussed previously in Section~\ref{sec:rank}.  
%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
%that is preferred.  
In both cases, the national versus international comparison
produced stronger \eds{lower?} p values than the 2019 versus 2022 comparison, with the
national vs international \texttt{exact = TRUE} test producing a p value of 
$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
is a highly significant test result and shows that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were a few months prior.  For the 2019 versus 2022
same athlete comparison with \texttt{exact = TRUE} argument, the p value was
significant at $\alpha = 0.01$, however when \texttt{exact = FALSE}, the p
value was higher and only significant at $\alpha = 0.05$. 
\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
Nonetheless, these
tests help to show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed previously.

\begin{table}
  \centering
  \caption{}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic \\ 
   \midrule
   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10^{-2}$ \\
   2022 vs 2023 & $3.0\cdot10^{-3}$ & $1.0\cdot10^{-2}$ \\
   Nat vs Inter & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. Nat vs Inter compares data from the same athletes who competed
  at the 2022 national-level championships and the 2022 World Track and Field
  Championships.}
  \label{tab:Clusrankresults}
\end{table}

@@

This paper does not consider data from the women's 100 meter hurdle, and although
that was initially because of concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 was at best only
slightly faster relative to other championships going back to 2003 
%\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
%the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
%results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
%womens 100m hurdles.} 
Likewise, the rank-based comparisons resulted in p values
that were not significant. 
\eds{could that be because of even smaller sample sizes?}
  The lack of evidence from the women's data makes it 
harder to conclude that the sensors used at the 2022 championships were faulty, 
although it is unknown if the sensors for the men's 110 meter hurdle and women's
100 meter hurdle were the same.  
\eds{is any calibration of the sensors needed (e.g. for weight?) that might
also play a role for women vs men. (My husband asked me about this as a possible 
reason for the difference in results for men and women, but he did not know
if any calibration is actually done.)}
\of{I honestly am not sure, do you think that this is worth investigating, or
beyond the scope of this paper? I am not sure if the mechanics of Seiko's
technology is public knowledge.}
